---
title: "resampling"
author: "AnyelaCamargo"
date: "10 March 2018"
output: html_document
---

```{r, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
library('ggplot2')
library('dismo')
library('rgdal')
library('foreign')
library('sp')
library('dplyr')
library('tmap')
library('ncdf4')
library('gstat')

```
The following examples make use of a NetCDF file of climate data from the Climate Research Unit http://www.cru.uea.ac.uk/data, consisting of long-term mean values (1961-1990) of near-surface air temperature on a 0.5-degree grid (for land points). The dimensions of the array are 720 (longitudes) x 360 (latitudes) x 12 (months).

The idea of the example is to show how weather variables can be interpolated (or extrapolated). 

##Load nc file and print main settings
```{r, include=FALSE}
fname = 'cru10min30_tmp.nc'
ncin <- nc_open(fname)
raster(fname) # 
ncin

plot(raster(fname))
```

##Get lat, lon, time and weather variables


```{r, include=TRUE}
st <- list()
st$lon <- ncvar_get(ncin,"lon")
st$lat <- ncvar_get(ncin,"lat")
st$time <- ncvar_get(ncin, "time")
st$tunits <- ncatt_get(ncin, "time", "units")
st$attrib_names <- names(ncin$var)
st$mnumber = length(time)
st$tp <- ncvar_get(ncin, st$attrib_names[2]) # get tmp array
```

## Convert time values
The time variable comes in the hour format. To transform it to the yy-mm-dd hh:mm format use the following command:


```{r, include=TRUE}

    as.POSIXct(st$time*3600, origin='1900-01-01 00:00')

```

## Generate grid with original data

Here tmp data corresponding to the first time point st$time[1] (e.g. 27773.5)) will be regrid/resample it to a smaller resolution. First extract subset to regrid. Remember, is just the number of hours since, you'd need to transform it to a different format to make sense of the number. 

```{r, include=TRUE}
 grd <- expand.grid(st$lon, st$lat, st$time)
    # Set names
    names(grd) <- c("lon", "lat", "time")
    # Put the data for qnet in a dataframe
    vdata <- data.frame(tp = as.numeric(st$tp)) %>% tbl_df()
    colnames(vdata) = st$attrib_names[2]
    # Bind with the grid
    subdata <- grd %>%
      tbl_df() %>%
      bind_cols(vdata) %>%
      filter(time == 27773.5) %>% # select time point 
      select_("-time") %>%
      na.omit()
```




## Convert to raster

```{r, include=TRUE}
coordinates(subdata) <- ~ lon + lat
  gridded(subdata) <- TRUE
  
plot(subdata)
```

## Regrid / Interpolate to UK only

```{r, include=TRUE}

  coor_list <- list()
 
  coor_list$lonmin <- -11
  coor_list$lonmax <- 2
  coor_list$latmin <- 48 
  coor_list$latmax <- 60
  
  res = 0.05
  data.grid <- expand.grid(lon = seq(coor_list$lonmin, coor_list$lonmax, res), 
                           lat = seq(coor_list$latmin, coor_list$latmax, res))
  coordinates(data.grid) <- ~ lon + lat
  gridded(data.grid) <- TRUE
   
  # IDW interpolation
  f <- as.formula(paste(st$attrib_name[2], ' ~ 1', sep=''))
  int_data <- idw(f, locations = subdata, newdata = data.grid, nmax = 5)

```

## Finally Plot the new grid

```{r, include=TRUE}

plot(int_data)

r <- raster(int_data)
```

## Convert to raster and show new grid
```{r, include=TRUE}
head(coordinates(r))
```
```
